The Effects of weather on maize yield: New evidence from Kenya

### 2018-12-25 writing/draft3/draft18.tex
write and editing...

Besides cumulative seasonal precipitation and average temperature, CV of both temperature and precipitation and also length and number of dry spells were significant (see the first column of table~\ref{MainEst}). As it is apparent from figures~\ref{MarEff2}(a) and \ref{MarEff2}(b) in the appendix, the relationship of maize yield and each of these variables is linear. In accordance with our expectation, the yield decreases with increasing CV of precipitation. \hl{However, we obtained a peculiar result for CV of temperature as the yield seem to increase with increasing variability of temperature. This is probably because...}

As we can see in figures~\ref{MarEff3}(a) and \ref{MarEff3}(b) in the appendix, yield declines with increasing number of dry spells and with length of maximum dry spell linearly. This is in accordance with our expectations.

According to both magnitude of the standardised coefficients and F-values in the first two columns of table~\ref{MainEst}, the most important variable of the model for all counties is average seasonal temperature. The temperature appears to be much more important than the second most important variable in this model which is cumulative precipitation. The value of the F-statistic corresponding to temperature is $12.010$ while the F-statistic of the linear term of cumulative precipitation is equal to $5.358$ and the F-statistic of its quadratic term is $4.263$. Also, according to its lowest p-value ($6\times10^{-4}$), temperature is the most important covariate.

We will now discuss and compare the results of the models for the subsamples of the ASAL counties (the third and fourth columns of table~\ref{MainEst}) and the non-ASAL counties (the fifth and sixth columns of table~\ref{MainEst}).

@@

\begin{threeparttable}[H]
\singlespacing
\caption{Precipitation and temperature measures considered}\label{vars}
\begin{tabular}{@{}lp{15cm}}
\\[-1em]
\br
\\[-1em]
$\bullet$& Total rainfall over the rainy season\\
$\bullet$& Coefficient of variation of the rainfall during the rainy season\\
$\bullet$& Maximum length of dry spell during the rainy season (in number of days)\\
$\bullet$& Number of dry spells during the rainy season: a dry spell defined as $4$ consecutive days without rain or more\tnote{a}\\
$\bullet$& Number of dry spells during the rainy season: a dry spell defined as $10$ consecutive days without rain or more\tnote{a}\\
$\bullet$& Number of dry spells during the rainy season: a dry spell defined as $20$ consecutive days without rain or more\tnote{a}\\
$\bullet$& Average temperature during the rainy season\\
$\bullet$& Standard deviation of temperature during the rainy season\\
$\bullet$& Cumulative degree days during the rainy season (excluding the days when maximum temperature above $30^{\circ}$C or below $10^{\circ}$C\\
$\bullet$& Number of heatwave days during the rainy season when max. temperature $>35^{\circ}$C
\\
$\bullet$& Maximum daily rainfall  - to control for possible floods
\\
$\bullet$& Sum of precipitation amount on days where precipitation is above $90^{th}$ percentile of precip. of the whole period\tnote{b}\\
$\bullet$& Sum of precipitation amount on days where precipitation is above $95^{th}$ percentile of precip. of the whole period\tnote{c}\\
$\bullet$& Sum of precipitation amount on days where precipitation is above $99^{th}$ percentile of precip. of the whole period\tnote{d}\\
$\bullet$& Number of days where precipitation is above $90^{th}$ percentile of precip. of the whole period\tnote{b}\\
$\bullet$& Number of days where precipitation is above $95^{th}$ percentile of precip. of the whole period\tnote{c}\\
$\bullet$& Number of days where precipitation is above $99^{th}$ percentile of precip. of the whole period\tnote{d}\\
\br
\end{tabular}
\begin{tablenotes}
  \begin{footnotesize}
  \item[a] Threshold for a dry day was considered $1$mm.
  \item[b] $90^{th}$ percentile was calculated for the subsample of wet days, that is the days where precipitation is above or equal to $1$mm.
\item[c] $95^{th}$ percentile was calculated for the subsample of wet days, that is the days where precipitation is above or equal to $1$mm.
  \item[d] $99^{th}$ percentile was calculated for the subsample of wet days, that is the days where precipitation is above or equal to $1$mm.
  \end{footnotesize}
\end{tablenotes}
\end{threeparttable}

>>>>

Besides cumulative seasonal precipitation and average temperature, CV of both temperature and precipitation and also length and number of dry spells were significant (see the first column of table~\ref{MainEst}). As it is apparent from figures~\ref{MarEff2}(a) and \ref{MarEff2}(b) in the appendix, the relationship of maize yield and each of these variables is linear. In accordance with our expectation, the yield decreases with increasing CV of precipitation. \hl{However, the result which we found for CV of temperature is a bit peculiar as the yield seem to increase with increasing variability of temperature. This is probably because...}

As we can see in figures~\ref{MarEff3}(a) and \ref{MarEff3}(b) in the appendix, yield declines with increasing number of dry spells and with increasing length of maximum dry spell linearly. This is in accordance with our expectations.

According to the magnitude of the standardised coefficients and F-values in the first two columns of table~\ref{MainEst}, the most important variable is average seasonal temperature which appears to be much more important than the remaining predictors. The second most important variable is total seasonal precipitation. The value of the F-statistic corresponding to temperature is $12.010$ while the F-statistic of the linear term of cumulative precipitation is equal to $5.358$ and the F-statistic of its quadratic term is $4.263$. Similarly, according to the Satterthwaite's p-values, average temperature is the most important predictor as the corresponding Satterthwaite's p-value is equal to $6\times10^{-4}$ which is much lower than the p-values of the other explanatory variables.

We will now discuss and compare the results of the models for the subsamples of the ASAL counties (the third and fourth columns of table~\ref{MainEst}) and the non-ASAL counties (the fifth and sixth columns of table~\ref{MainEst}). As for the subsample of the ASAL counties, the only significant fixed effects included average seasonal temperature and length and number of dry spells, all of which were negative. Interestingly, different predictors were  found significant for the subsample of the non-ASAL counties. These included seasonal precipitation and its square (the first one was positive and the second one negative..\textcolor {red}{here}. )

@@

\begin{threeparttable}[H]
\singlespacing
\caption{Precipitation and temperature measures considered}\label{vars}
\begin{tabular}{@{}lp{15cm}}
\\[-1em]
\br
\\[-1em]
$\bullet$& Total precipitation over the rainy season\\
$\bullet$& Coefficient of variation of the precipitation during the rainy season\\
$\bullet$& Maximum length of dry spell during the rainy season (in number of days)\\
$\bullet$& Number of dry spells during the rainy season: a dry spell defined as $4$ consecutive days without rain or more\tnote{a}\\
$\bullet$& Number of dry spells during the rainy season: a dry spell defined as $10$ consecutive days without rain or more\tnote{a}\\
$\bullet$& Number of dry spells during the rainy season: a dry spell defined as $20$ consecutive days without rain or more\tnote{a}\\
$\bullet$& Average temperature during the rainy season\\
$\bullet$& Standard deviation of temperature during the rainy season\\
$\bullet$& Cumulative degree days during the rainy season (excluding the days when maximum temperature above $30^{\circ}$C or below $10^{\circ}$C\\
$\bullet$& Number of heatwave days during the rainy season when max. temperature $>35^{\circ}$C
\\
$\bullet$& Maximum daily precipitation  - to control for possible floods
\\
$\bullet$& Sum of precipitation amount on days where precipitation is above $90^{th}$ percentile of precip. of the whole period\tnote{b}\\
$\bullet$& Sum of precipitation amount on days where precipitation is above $95^{th}$ percentile of precip. of the whole period\tnote{c}\\
$\bullet$& Sum of precipitation amount on days where precipitation is above $99^{th}$ percentile of precip. of the whole period\tnote{d}\\
$\bullet$& Number of days where precipitation is above $90^{th}$ percentile of precip. of the whole period\tnote{b}\\
$\bullet$& Number of days where precipitation is above $95^{th}$ percentile of precip. of the whole period\tnote{c}\\
$\bullet$& Number of days where precipitation is above $99^{th}$ percentile of precip. of the whole period\tnote{d}\\
\br
\end{tabular}
\begin{tablenotes}
  \begin{footnotesize}
  \item[a] Threshold for a dry day was considered $1$mm.
  \item[b] $90^{th}$ percentile was calculated for the subsample of wet days, that is the days where precipitation is above or equal to $1$mm.
\item[c] $95^{th}$ percentile was calculated for the subsample of wet days, that is the days where precipitation is above or equal to $1$mm.
  \item[d] $99^{th}$ percentile was calculated for the subsample of wet days, that is the days where precipitation is above or equal to $1$mm.
  \end{footnotesize}
\end{tablenotes}
\end{threeparttable}
